---
title: "Active Learning for Image Classification"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=FALSE, message=FALSE, warning=FALSE, fig.height=7.5)
rxOptions(reportProgress=0)
```

# Classifying wood knots

This is a followup to our earlier blog post "[Featurizing images: the shallow end of deep learning](http://blog.revolutionanalytics.com/2017/09/wood-knots.html#more)". That article contains the code for generating the features for the training and test datasets, which were saved to a csv file. Here we begin by loading that file.


```{r parameters}

### Libraries ###
library(dplyr)
library(tidyr)
library(ggplot2)
library(reticulate)
library(pROC)

### Meta-hyperparameters ###
set.seed(3)  ###

L1_PENALTY <- 1e-2
L2_PENALTY <- 1e-2

INITIAL_EXAMPLES_PER_CLASS <- 6  # cases from the labelled dataset used to train the initial model
ADDITIONAL_CASES_TO_LABEL <- 12  # cases to label per iteration
NUM_ITERATIONS <- 50
MONTE_CARLO_SAMPLES <- 100

# This order determines the order of factor levels
KNOT_CLASSES <- c("sound_knot", "dry_knot", "encased_knot")

LABELLED_FEATURIZED_DATA <- "data/labelled_knots_featurized_resnet18.Rds"
UNLABELLED_FEATURIZED_DATA <- "data/unlabelled_knots_featurized_large_resnet18.Rds"

unlabelled_knot_data_df <- readRDS(UNLABELLED_FEATURIZED_DATA)
labelled_knot_data_df <- readRDS(LABELLED_FEATURIZED_DATA)

# We'll pretend these come from our labellers
PSEUDOLABELS_FILE <- "data/unlabelled_knot_info_area_small.csv" 

inputs <- grep("^Feature", names(labelled_knot_data_df), value=TRUE)
FORM <- formula(paste0('~', paste(inputs, collapse=" + "), " - 1"))
# outcome <- "knot_class"
# FORM <- formula(paste0(outcome, '~', paste(inputs, collapse=" + "), " - 1"))

```


### Split labelled data into training and test sets

```{r split_train_and_test_sets}

class_samples <- lapply(KNOT_CLASSES, function(kc) sample(which(labelled_knot_data_df$knot_class == kc), INITIAL_EXAMPLES_PER_CLASS))

in_training_set <- (1:nrow(labelled_knot_data_df) %in% unlist(class_samples))

initial_training_set <- labelled_knot_data_df[in_training_set,]
TEST_SET <- labelled_knot_data_df[!in_training_set,]

table(initial_training_set$knot_class)

table(TEST_SET$knot_class)
table(TEST_SET$knot_class)/nrow(TEST_SET)
```


## Initial model for knot classes

First we build a model on the available training data, and test on the test data. This model will focus on classifying a knot image into three categories: "sound_knot", "dry_knot", and "encased_knot". In the blog post there was also a category for "decayed_knot", but we will consider decay a separate attribute, and not use it in this model. We will use only a small number of the available labelled cases for training, and the rest for testing. This makes a larger test set and a smaller training set than in the blog post.

### Fit model to initial training set

```{r tune_and_train_model}
source("woodknots_active_learning_lib.R")

pseudolabel_function <- get_pseudolabelling_function(PSEUDOLABELS_FILE, KNOT_CLASSES)

initial_model_results <- fit_and_evaluate_model(initial_training_set)

```

## Results for initial model


### ROC curves

```{r roc_curves}
mapply(plot, x=initial_model_results$roc_list, main=names(initial_model_results$roc_list), print.auc=TRUE) %>% invisible
```

#### Confusion matrix

```{r initial_model_confusion}

initial_model_results$confusion

```

### Performance summary

```{r initial_model_performance}
initial_model_results$performance

```

### Histograms of class scores

```{r class_score_histograms}

plot_class_histograms(initial_model_results$test_predictions)

```

### Plot of test cases on entropy surface

Here we'll stick to a 2D representation, where the yellow background indicates regions of lower entropy.

```{r plot_initial_class_separation}

plot_class_separation(initial_model_results$test_predictions)

```

## Iterate modelling, case selection, and (pseudo) labelling

These are the cases selected by the initial model for labelling:

```{r initial_model_results_selected}

initial_model_results$selected <- select_cases(initial_model_results$model, unlabelled_knot_data_df)

initial_model_results$selected

```

```{r iterate}

new_sample <- initial_model_results$selected %>% pseudolabel_function %>% get_new_pseudolabelled_sample(unlabelled_knot_data_df)

current_training_set <- rbind(initial_training_set, new_sample[names(initial_training_set)])

ALREADY_EVALUATED <- initial_model_results$selected$path

iteration_results <- lapply(1:NUM_ITERATIONS, function(i){
  results <- fit_and_evaluate_model(current_training_set)
  
  candidate_cases <- unlabelled_knot_data_df[(unlabelled_knot_data_df$path %in% setdiff(unlabelled_knot_data_df$path, ALREADY_EVALUATED)),]
  results$selected <- select_cases(results$model, candidate_cases)

  ALREADY_EVALUATED <<- c(ALREADY_EVALUATED, results$selected$path)
  results$selected_labelled <- results$selected %>% pseudolabel_function
  next_sample <- results$selected %>% pseudolabel_function %>% get_new_pseudolabelled_sample(unlabelled_knot_data_df)
  
  current_training_set <<- rbind(current_training_set, next_sample[names(current_training_set)])

  results
})
```


```{r mean_entropy_of_selected_cases_by_iteration}
mean_entropy <- sapply(iteration_results, function(ires) mean(ires$selected$entropy))
plot(mean_entropy, type='l', main="mean entropy of selected cases by iteration")
```

These are the cases selected at each iteration, together with the scores produced by the model for that iteration. The `knot_class` column was added by the pseudolabelling function.

```{r iteration_results_selected}
# lapply(iteration_results, function(ires) ires$selected_labelled)
```

This shows the change in the metrics, with each row showing an iteration. The 'negentropy' metric is the negative entropy across all three class probabilities.

```{r visualize_metrics_by_iteration}
do.call("rbind", lapply(iteration_results, function(ires) ires$performance))

```

### Visualizing improvement for actively learned model

Here we plot a series of ROC curves showing how performance changes with iterations of active learning.

```{r visualizing_improvement}

plot_roc_history("sound", initial_model_results, iteration_results)
plot_roc_history("dry", initial_model_results, iteration_results)
plot_roc_history("encased", initial_model_results, iteration_results)
```


### Final model results

```{r final_model}
final_model_results <- iteration_results[[NUM_ITERATIONS]]
```

### Confusion Matrix

```{r final_model_confusion_matrix}
final_model_results$confusion
```

### Performance summary

Summary of performance using cases selected with active learning:

```{r summary_of_preformance_using_selected_cases}

(selected_sample_results <- final_model_results$performance)
```

### Histograms of class scores for final model

```{r final_class_score_histograms}

plot_class_histograms(final_model_results$test_predictions)

```

### Entropy surface plot

```{r classifier_evolution}
plot_class_separation(final_model_results$test_predictions, main="Final")
```
